---
title: "Simple BUGS Model Examples"
author: "Bob"
date: "2025-09-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(dplyr)
```

# Simple statistical estimation example

In this example (provided by ChatGPT) we are going to simulate some random data with a given mean and standard deviation, then use JAGS to estimate these parameters.

```{r chatGPT_example}
# 1. Load jagsUI
library(jagsUI)

# 2. Simulate some data
set.seed(123)
n <- 20
y <- rnorm(n, mean = 5, sd = 2)

# 3. Prepare data list for JAGS
data_list <- list(
  y = y,
  n = n
)

# 4. Write JAGS model as a string
model_string <- "
model {
  # Likelihood
  for (i in 1:n) {
    y[i] ~ dnorm(mu, tau)
  }

  # Priors
  mu ~ dnorm(0, 0.001)      # vague prior for mean
  sigma ~ dunif(0, 10)      # uniform prior for SD
  tau <- 1 / (sigma^2)      # precision
}
"

# Write model to a temporary file
writeLines(model_string, con = "model.txt")

# 5. Set initial values
inits <- function() {
  list(mu = rnorm(1, 0, 1),
       sigma = runif(1, 0.1, 5))
}

# 6. Run the model with jagsUI
results <- jags(
  data = data_list,
  inits = inits,
  parameters.to.save = c("mu", "sigma"),
  model.file = "model.txt",
  n.chains = 3,
  n.iter = 5000,
  n.burnin = 1000,
  n.thin = 2
)

# 7. Summarize results
print(results)

# Optional: plot posterior distributions
plot(results)

# 7. Summarize results
print(results)

# Optional: plot posterior distributions
plot(results)
```

# Snake Weight

From [Linear regression models using Bayesian analysis in JAGS](https://biometry.github.io/APES//LectureNotes/StatsCafe/Linear_models_jags.html). I modified the code slightly to use jagsUI.

## Simple linear regression

```{r sim_snakes_1}

set.seed(42) # Set a random seed for reproducibility of the simulation

samplesize <- 30 # Number of data points
b_length <- sort(rnorm(samplesize)) # Body length (explanatory variable)

int_true <- 30 # True intercept
slope_true <- 10 # True slope
mu <- int_true + slope_true * b_length # True means of normal distributions
sigma <- 5 # True standard deviation of normal distributions

b_mass <- rnorm(samplesize, mean = mu, sd = sigma) # Body mass (response variable)

snakes1 <- data.frame(b_length = b_length, b_mass = b_mass)
head(snakes1)
with(snakes1, plot(b_length, b_mass))

```

```{r snakes_jags-1}

jags_data_s1 <- with(snakes1, list(b_mass = b_mass, b_length = b_length, N = length(b_mass)))

init_values <- function(){
	list(alpha = rnorm(1), beta = rnorm(1), sigma = runif(1))
}

snake_model_code <- "
model{
	# Likelihood:
	for (i in 1:N){
		b_mass[i] ~ dnorm(mu[i], tau) # tau is precision (1 / variance)
		mu[i] <- alpha + beta * b_length[i]
	}
	# Priors:
	alpha ~ dnorm(0, 0.01) # intercept
	beta ~ dnorm(0, 0.01) # slope
	sigma ~ dunif(0, 100) # standard deviation
	tau <- 1 / (sigma * sigma) # sigma^2 doesn't work in JAGS
}
" %>% textConnection()


snake_params <- c("alpha", "beta", "sigma")

fit_lm1 <- jags(data = jags_data_s1, 
											inits = init_values, 
											parameters.to.save = snake_params, 
											model.file = snake_model_code,
											n.chains = 3, 
											n.iter = 12000, 
											n.burnin = 2000, 
											n.thin = 10, 
											DIC = F)

```
```{r snake_results}

plot(fit_lm1)

```

### plot credible interval as in the article.

```{r plot_credible_interval}

nvalues <- 100
b_length_new <- seq(min(snakes1$b_length), max(snakes1$b_length), length.out = nvalues)

lm1_mcmc_combi <- do.call("rbind", fit_lm1$samples)

pred_mean_mean <- mean(lm1_mcmc_combi[, "alpha"]) + b_length_new * mean(lm1_mcmc_combi[, "beta"])

pred_mean_dist <- matrix(NA, nrow = nrow(lm1_mcmc_combi), ncol = nvalues)
for (i in 1:nrow(pred_mean_dist)){
	pred_mean_dist[i,] <- lm1_mcmc_combi[i,"alpha"] + b_length_new * lm1_mcmc_combi[i,"beta"]
}
credible_lower <- apply(pred_mean_dist, MARGIN = 2, quantile, prob = 0.025)
credible_upper <- apply(pred_mean_dist, MARGIN = 2, quantile, prob = 0.975)

lm1_mcmc_combi_rep <- do.call(rbind, rep(list(lm1_mcmc_combi), 50)) # replication

# Draw random values for all parameter combinations (rows) and body length values (columns):
pred_data_dist <- matrix(NA, nrow = nrow(lm1_mcmc_combi_rep), ncol = nvalues)
for (i in 1:nrow(pred_data_dist)){
	pred_data_dist[i,] <- lm1_mcmc_combi_rep[i,"alpha"] + b_length_new * lm1_mcmc_combi_rep[i,"beta"] +
		rnorm(nvalues, mean = 0, sd = lm1_mcmc_combi_rep[i, "sigma"])
}

# Calculate quantiles:
uncertain_lower <- apply(pred_data_dist, MARGIN = 2, quantile, prob = 0.025)
uncertain_upper <- apply(pred_data_dist, MARGIN = 2, quantile, prob = 0.975)

plot(b_mass ~ b_length, data = snakes1)
lines(b_length_new, pred_mean_mean)
lines(b_length_new, credible_lower, lty = 2)
lines(b_length_new, credible_upper, lty = 2)
lines(b_length_new, uncertain_lower, lty = 2, col = "red")
lines(b_length_new, uncertain_upper, lty = 2, col = "red")

```


## Same model with an additional categorical predictor

```{r sim_snakes_2}
set.seed(42)

samplesize <- 50 # Larger sample size because we're fitting a more complex model
b_length <- sort(rnorm(samplesize)) # Body length
sex <- sample(c(0, 1), size = samplesize, replace = T) # Sex (0: female, 1: male)

int_true_f <- 30 # Intercept of females
int_true_m_diff <- 5 # Difference between intercepts of males and females
slope_true_f <- 10 # Slope of females
slope_true_m_diff <- -3 # Difference between slopes of males and females

mu <- int_true_f + sex * int_true_m_diff + (slope_true_f + sex * slope_true_m_diff) * b_length # True means
sigma <- 5 # True standard deviation of normal distributions

b_mass <- rnorm(samplesize, mean = mu, sd = sigma) # Body mass (response variable)

# Combine into a data frame:
snakes2 <- data.frame(b_length = b_length, b_mass = b_mass, sex = sex)
head(snakes2)

plot(b_mass ~ b_length, 
		 col = (sex + 1), 
		 data = snakes2, 
		 x_lab="relative length",
		 y_lab="body mass")
text(2, 10, "F", col=1)
text(2, 7, "M", col=2)
```

```{r snakes_model_2}
jagsdata_s2 <- with(snakes2, list(b_mass = b_mass, b_length = b_length, sex = sex, N = length(b_mass)))

lm2_jags <- "
model{
	# Likelihood:
	for (i in 1:N){
		b_mass[i] ~ dnorm(mu[i], tau) # tau is precision (1 / variance)
		mu[i] <- alpha[1] + sex[i] * alpha[2] + (beta[1] + beta[2] * sex[i]) * b_length[i]
	}
	# Priors:
	for (i in 1:2){
		alpha[i] ~ dnorm(0, 0.01)
		beta[i] ~ dnorm(0, 0.01)
	}
	sigma ~ dunif(0, 100)
	tau <- 1 / (sigma * sigma)
}
" %>% textConnection()

init_values <- function(){
	list(alpha = rnorm(2), beta = rnorm(2), sigma = runif(1))
}

params <- c("alpha", "beta", "sigma")

fit_lm2 <- jags(data = jagsdata_s2, 
								inits = init_values, 
								parameters.to.save = params, 
								model.file = lm2_jags,
								n.chains = 3, 
								n.iter = 12000, 
								n.burnin = 2000, 
								n.thin = 10, 
								DIC = F)

plot(fit_lm2)
```


## A linear mixed model

```{r linear_mixed_model}
samplesize <- 200
nsites <- 10 # Number of sites
b_length <- sort(rnorm(samplesize)) # Body length (explanatory variable)
sites <- sample(1:10, samplesize, replace = T) # Site (grouping variable)
table(sites)

int_true_mean <- 45 # True mean intercept
int_true_sigma <- 10 # True standard deviation of intercepts
int_true_sites <- rnorm(n = nsites, mean = int_true_mean, sd = int_true_sigma) # True intercept of each site

# Intercept of each snake individual (depending on the site where it was captured):
sitemat <- matrix(0, nrow = samplesize, ncol = nsites)
for (i in 1:nrow(sitemat)) sitemat[i, sites[i]] <- 1
int_true <- sitemat %*% int_true_sites

slope_true <- 10 # True slope
mu <- int_true + slope_true * b_length # True means of normal distributions
sigma <- 5 # True standard deviation of normal distributions

b_mass <- rnorm(samplesize, mean = mu, sd = sigma) # Body mass (response variable)

snakes3 <- data.frame(b_length = b_length, b_mass = b_mass, site = sites)
head(snakes3)

plot(b_mass ~ b_length, col = site, data = snakes3)

```
```{r fit_mixed_model}
Nsites <- length(levels(as.factor(snakes3$site)))

jagsdata_s3 <- with(snakes3, list(
	b_mass = b_mass, 
	b_length = b_length, 
	site = site,
	N = length(b_mass), Nsites = Nsites)
)

lm3_jags <- "model{
	# Likelihood:
	for (i in 1:N){
		b_mass[i] ~ dnorm(mu[i], tau) # tau is precision (1 / variance)
		mu[i] <- alpha + a[site[i]] + beta * b_length[i] # Random intercept for site
	}
	# Priors:
	alpha ~ dnorm(0, 0.01) # intercept
	sigma_a ~ dunif(0, 100) # standard deviation of random effect (variance between sites)
	tau_a <- 1 / (sigma_a * sigma_a) # convert to precision
	for (j in 1:Nsites){
		a[j] ~ dnorm(0, tau_a) # random intercept for each site
	}
	beta ~ dnorm(0, 0.01) # slope
	sigma ~ dunif(0, 100) # standard deviation of fixed effect (variance within sites)
	tau <- 1 / (sigma * sigma) # convert to precision
}" %>% textConnection()

init_values <- function(){
	list(alpha = rnorm(1), sigma_a = runif(1), beta = rnorm(1), sigma = runif(1))
}

params <- c("alpha", "beta", "sigma", "sigma_a")

fit_lm3 <- jags(data = jagsdata_s3, inits = init_values, parameters.to.save = params, model.file = lm3_jags,
			   n.chains = 3, n.iter = 20000, n.burnin = 5000, n.thin = 10, DIC = F)

fit_lm3

```
